Automated Formal Synthesis of Digital Controllers for State-Space Physical Plants

### 2017-01-25 cav2017/cav2017.tex
edits

The models we consider have two sources of error due to numeric representation. 
The first is the numeric error introduced by the fixed-point numbers employed to model the plant, i.e., to represent the plant dynamics $A_d$, $B_d$ and $x_k$. 
The second is the quantization error introduced by the digital controller, 
which performs operations in the domain of fixed-point numbers. 
In this section we outline the notation for the fixed-point representation of numbers, and briefly describe the errors introduced. A formal outline is in Appendix~\ref{appendix:numerical_errors}.

@@

Let us first consider the effect of the quantization errors on safety. 
Within the digital controller block, 
state values are manipulated at low digital precision, 
alongside the vector multiplication $Kx$.   
%computed considering the error introduced by the digital controller caused by performing the vector multiplication $Kx$ at low digital precision.
As such, the inputs are computed using the following equation: 
\begin{align*}
u_{k}&=-(\mathcal{F}_{\langle I_c,F_c \rangle}(K)\cdot\mathcal{F}_{\langle I_c,F_c \rangle}(x_{k})). 
\end{align*}

@@

An additional error is caused by the representation of the plant dynamics, namely 
\begin{align*}
x_{k+1} &=\mathcal{F}_{\langle I_p,F_p \rangle}(A_d) \mathcal{F}_{\langle I_p,F_p \rangle}(x_{k}) + \mathcal{F}_{\langle I_p,F_p \rangle}(B_d)\mathcal{F}_{\langle I_p,F_p \rangle}(u_{k}).
\end{align*}

@@

All benchmarks were discretized with different sample times using 
the approach described in the Appendix~\ref{sec:appendix:LTIbackground}. 
All experiments were performed considering $\underline{x_{i}}=-1$ and 
$\overline{x_{i}}=1$ and the inputs $u_{k}=0, \forall k>0$.
We have conducted the experimental evaluation on a 12-core 2.40 GHz
Intel Xeon E5-2440 with 96 GB of RAM and Linux OS. We used the Linux \emph{times} command
to measure CPU time used for each benchmark. The experimental runtime was limited to
one hour per benchmark.
%

@@

So $\delta_T$ is the truncation error and it will propagate across operations.
\item {\bf Rounding:}
The following errors appear in simple operations. Let $c_1, c_2$ and $c_3$ be real numbers, and $\delta_{T1}$ and $\delta_{T2}$ be the truncation errors caused by representing $c_1$ and $c_2$ in the fixed-point domain as above.
\begin{enumerate}
\item Addition/Subtraction: these operations only propagate errors coming from truncation of the operands, namely 
$\mathcal{F}_{\langle I,F \rangle}(c_1) \pm \mathcal{F}_{\langle I,F \rangle}(c_2) = c_3 + \delta_3$ with $|\delta_3| \leq |\delta_{T1}| + |\delta_{T2}|$.
\item Multiplication: $\mathcal{F}_{\langle I,F \rangle}(c_1) * \mathcal{F}_{\langle I,F \rangle}(c_2) =  c_3 + \delta_3$ with $|\delta_3| \leq |\delta_{T1}*\mathcal{F}_{\langle I,F \rangle}(c_2)| + |\delta_{T2}*\mathcal{F}_{\langle I,F \rangle}(c_1)| + c_m$, where $c_m=2^{-F}$ as above.
\item Division: the operations performed by our controllers in the FWL domain do not include division. However, we do use division in computations at the precision of the plant. Here the error depends on whether the divisor is greater or smaller than the dividend. 
$\mathcal{F}_{\langle I,F \rangle}(c_1) / \mathcal{F}_{\langle I,F \rangle}(c_2) =  c_3 + \delta_{T3}$ where $\delta_{T3}$ is $(\delta_{T2}*c_1 - \delta_{T1}*c_2)/(\delta_{T2}^2 - \delta_{T2} c_2)$,
%$\mathcal{F}_{\langle I,F \rangle}(c_1) / \mathcal{F}_{\langle I,F \rangle}(c_2) =  (c_1 - \delta_{T1})/ (c_2 + \delta_{T2}) + c_m $, where $c_m=2^{-F}$ as above.
\end{enumerate}

>>>>

The models we consider have two sources of error due to numeric representation. 
The first is the numeric error introduced by the fixed-point numbers employed to model the plant, i.e., to represent the plant dynamics $A_d$, $B_d$ and $x_k$. 
The second is the quantization error introduced by the digital controller, 
which performs operations on fixed-point numbers. 
In this section we outline the notation for the fixed-point representation of numbers, and briefly describe the errors introduced. A formal outline is in Appendix~\ref{appendix:numerical_errors}.

@@

Let us first consider the effect of the quantization errors on safety. 
Within the controller, 
state values are manipulated at low precision, 
alongside the vector multiplication $Kx$.   
%computed considering the error introduced by the digital controller caused by performing the vector multiplication $Kx$ at low digital precision.
The inputs are computed using the following equation: 
%
\begin{align*}
u_{k}&=-(\mathcal{F}_{\langle I_c,F_c \rangle}(K)\cdot\mathcal{F}_{\langle I_c,F_c \rangle}(x_{k})). 
\end{align*}

@@

An additional error is caused by the representation of the plant dynamics, namely 
%
\begin{align*}
x_{k+1} &=\mathcal{F}_{\langle I_p,F_p \rangle}(A_d) \mathcal{F}_{\langle I_p,F_p \rangle}(x_{k}) + \mathcal{F}_{\langle I_p,F_p \rangle}(B_d)\mathcal{F}_{\langle I_p,F_p \rangle}(u_{k}).
\end{align*}

@@

All benchmarks were discretized with different sample times using 
the approach described in the Appendix~\ref{sec:appendix:LTIbackground}. 
All experiments were performed considering $\underline{x_{i}}=-1$ and 
$\overline{x_{i}}=1$ and the inputs $u_{k}=0, \forall k>0$.
We have conducted the experimental evaluation on a 12-core 2.40\,GHz
Intel Xeon E5-2440 with 96\,GB of RAM and Linux OS. We used the Linux \emph{times} command
to measure CPU time used for each benchmark. The experimental runtime was limited to
one hour per benchmark.
%

@@

So $\delta_T$ is the truncation error and it will propagate across operations.
\item {\bf Rounding:}
The following errors appear in basic operations. Let $c_1, c_2$ and $c_3$ be real numbers, and $\delta_{T1}$ and $\delta_{T2}$ be the truncation errors caused by representing $c_1$ and $c_2$ in the fixed-point domain as above.
\begin{enumerate}
\item Addition/Subtraction: these operations only propagate errors coming from truncation of the operands, namely 
$\mathcal{F}_{\langle I,F \rangle}(c_1) \pm \mathcal{F}_{\langle I,F \rangle}(c_2) = c_3 + \delta_3$ with $|\delta_3| \leq |\delta_{T1}| + |\delta_{T2}|$.
\item Multiplication: $\mathcal{F}_{\langle I,F \rangle}(c_1) * \mathcal{F}_{\langle I,F \rangle}(c_2) =  c_3 + \delta_3$ with $|\delta_3| \leq |\delta_{T1}*\mathcal{F}_{\langle I,F \rangle}(c_2)| + |\delta_{T2}*\mathcal{F}_{\langle I,F \rangle}(c_1)| + c_m$, where $c_m=2^{-F}$ as above.
\item Division: the operations performed by our controllers in the FWL domain do not include division. However, we do use division in computations at the precision of the plant. Here the error depends on whether the divisor is greater or smaller than the dividend. 
$\mathcal{F}_{\langle I,F \rangle}(c_1) / \mathcal{F}_{\langle I,F \rangle}(c_2) =  c_3 + \delta_{T3}$ where $\delta_{T3}$ is $(\delta_{T2}*c_1 - \delta_{T1}*c_2)/(\delta_{T2}^2 - \delta_{T2} c_2)$,
%$\mathcal{F}_{\langle I,F \rangle}(c_1) / \mathcal{F}_{\langle I,F \rangle}(c_2) =  (c_1 - \delta_{T1})/ (c_2 + \delta_{T2}) + c_m $, where $c_m=2^{-F}$ as above.
\end{enumerate}
